Prediction Assignment
=====================

Assefa A.

### Executive summary
The goal of this project is to discriminate between different weight lifting activities. The dataset for modeling was first dividing in to training (70% of the total data) and test (30%) datasets. A classification model was build considering taining set, crossvalidation, and randomforest algorithm. The acuracy of the model was evaluated using the test dataset. Finally, the model was used to predict a new separate dataset.

#### Downloading and cleaning dataset

```{r, cache = TRUE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
              "modelData.csv", method="curl")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
              "predictData.csv", method="curl")
modelData <- read.csv("modelData.csv")
predictData <- read.csv("predictData.csv")
modelData$classe <- as.factor(modelData$classe)

dim(modelData)
summary(modelData$classe)
```

```{r}
naData <- sapply(modelData, function (x) any(is.na(x) | x == ""))
nnaIndex <- !naData & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(naData))

cleanIndex <- c("classe", names(naData)[nnaIndex])
cleanData <- modelData[, cleanIndex]

dim(cleanData)
summary(cleanData$classe)
```

#### Dividing in to trainng and test datasets
```{r}
library(caret)

trainIndex <- createDataPartition(y=cleanData$classe, p=0.7, list=FALSE)
trainData <- cleanData[trainIndex, ]
testData <- cleanData[-trainIndex, ]
summary(trainData$classe)
summary(testData$classe)
```

#### Prediction model using cross validation
On testing the model using the test dataset, the model is 99.5% accurate.
```{r}
library(randomForest)

modelControl <- trainControl(method = "repeatedcv", number = 10, repeats = 10)
modelTrain <- randomForest(classe~., data=trainData, trControl = modelControl)

mean(predict(modelTrain, testData) == testData$classe)*100
```

#### Predicting the class of the new independent dataset and exporting the results as .txt files
```{r}
pred <- predict(modelTrain, newdata = predictData)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pred)
```